---
title: "Problem Set 1"
author: "Yifan Li"
format: pdf
editor: visual
---

## Problem 1

# a

Using `read.table` to read a file written in txt format. For the separation, using ','. Then according to the description file, 'wine.names', there are 14 attributes in the data file with a class number listed in the first column. So adding `col.names` in the code `read.table`. Such that, one can produce a `data.frame` object with appropriate columns names.

```{r}
wines_data <- read.table("wine.data",
                         sep = ",",
                         col.names=c('class_number',
                                     'Alcohol',
                                     'Malic_acid',
                                     'Ash',
                                     'Alcalinity_of_ash',
                                     'Magnesium',
                                     'Total_phenols',
                                     'Flavanoids',
                                     'Nonflavanoid_phenols',
                                     'Proanthocyanins',
                                     'Color_intensity',
                                     'Hue',
                                     'OD280_OD315_of_diluted_wines',
                                     'Proline'))
```

# b

First, using `wines_data['class_number]==i` for `i in [1, 2, 3]` to create a new `data.frame` that has `True` only if the class numbers match with `i`. After that, using a `sum` function to compute the number of `True`, which is the number of the wine class.

```{r}
num_class_one <- sum(wines_data['class_number'] == 1) 
num_class_two <- sum(wines_data['class_number'] == 2)
num_class_three <- sum(wines_data['class_number'] == 3)
```

The results are:

```{r}
num_class_one
```

```{r}
num_class_two
```

```{r}
num_class_three
```

So, the number of wines within each class is correct as reported in the file “wine.names”.

# c

1.  The correlation between alcohol content and color intensity can be derived from a function `cor`. The alcohol content has variable name `Alcohol`, the color intensity has variable name `Color_intensity`. So the input of the function will be:

```{r}
cor(wines_data['Alcohol'], wines_data['Color_intensity'])
```

2.  For each class, first the whole data from that class will be extracted, then the correlation between alcohol content and color intensity will be calculated.

For class one:

```{r}
class_one <- wines_data[wines_data['class_number'] == 1, ]
class_one_cor <- cor(class_one['Alcohol'], class_one['Color_intensity'])
class_one_cor
```

For class two:

```{r}
class_two <- wines_data[wines_data['class_number'] == 2, ]
class_two_cor <- cor(class_two['Alcohol'], class_two['Color_intensity'])
class_two_cor
```

For class three:

```{r}
class_three <- wines_data[wines_data['class_number'] == 3, ]
class_three_cor <- cor(class_three['Alcohol'], class_three['Color_intensity'])
class_three_cor
```

Through comparison, one will find that class one has the highest correlation which is `0.4082913`, while class two has the lowest correlation which is `0.2697891`.

3.  To find the wine with highest color intensity, using `which.max` function, with attributes `wines_data$Color_intensity`. This will yield the index of the wine with highest color intensity. Then using this index to find the wine, after that extract its alcohol content.

```{r}
index <- which.max(wines_data$Color_intensity)
target_wine <- wines_data[index, ]
target_wine$Alcohol
```

Finally extract the alcohol content from the target wine, which is `14.34`.

4.  First, find the number of wines that have a higher content of proanthocyanins than ash. Then divide it by the sum of three classes of wines, which will give us the percentage of wines had a higher content of proanthocyanins compare to ash, which is `8.426966%`.

```{r}
num <- sum(wines_data$'Proanthocyanins' > wines_data$'Ash')
percentage <- num * 100 / (num_class_one + num_class_two + num_class_three)
percentage
```

# d

```{r}
average_table <- data.frame(id = 1: 4,
                            class_number = c('overall', '1', '2', '3'),
                            
                            Mean_Alcohol = c(mean(wines_data$Alcohol), 
                                             mean(class_one$Alcohol), 
                                             mean(class_two$Alcohol), 
                                             mean(class_three$Alcohol)),
                            
                            Mean_Malic_acid = c(mean(wines_data$Malic_acid),
                                                mean(class_one$Malic_acid), 
                                                mean(class_two$Malic_acid),
                                                mean(class_three$Malic_acid)),
                            
                            Mean_Ash = c(mean(wines_data$Ash),
                                         mean(class_one$Ash),
                                         mean(class_two$Ash),
                                         mean(class_three$Ash)),
                            
                            Mean_Alcalinity_of_ash = c(
                              mean(wines_data$Alcalinity_of_ash),
                              mean(class_one$Alcalinity_of_ash),
                              mean(class_two$Alcalinity_of_ash),
                              mean(class_three$Alcalinity_of_ash)),
                            
                            Mean_Magnesium = c(mean(wines_data$Magnesium),
                                               mean(class_one$Magnesium),
                                               mean(class_two$Magnesium),
                                               mean(class_three$Magnesium)),
                            
                            Mean_Total_phenols = c(mean(wines_data$Total_phenols),
                                                   mean(class_one$Total_phenols),
                                                   mean(class_two$Total_phenols),
                                                   mean(class_three$Total_phenols)),
                            
                            Mean_Flavanoids = c(mean(wines_data$Flavanoids),
                                                mean(class_one$Flavanoids),
                                                mean(class_two$Flavanoids),
                                                mean(class_three$Flavanoids)),
                            
                            Mean_Nonflavanoid_phenols = c(
                              mean(wines_data$Nonflavanoid_phenols),
                              mean(class_one$Nonflavanoid_phenols),
                              mean(class_two$Nonflavanoid_phenols),
                              mean(class_three$Nonflavanoid_phenols)),
                            
                            Mean_Proanthocyanins = c(mean(wines_data$Proanthocyanins),
                                                     mean(class_one$Proanthocyanins),
                                                     mean(class_two$Proanthocyanins),
                                                     mean(class_three$Proanthocyanins)),
                            
                            Mean_Color_intensity = c(mean(wines_data$Color_intensity),
                                                     mean(class_one$Color_intensity),
                                                     mean(class_two$Color_intensity),
                                                     mean(class_three$Color_intensity)),
                            
                            Mean_Hue = c(mean(wines_data$Hue), mean(class_one$Hue),
                                         mean(class_two$Hue), mean(class_three$Hue)),
                            
                            Mean_OD280_OD315_of_diluted_wines = c(
                                mean(wines_data$OD280_OD315_of_diluted_wines),
                                mean(class_one$OD280_OD315_of_diluted_wines),
                                mean(class_two$OD280_OD315_of_diluted_wines),
                                mean(class_three$OD280_OD315_of_diluted_wines)),
                            
                            Mean_Proline = c(mean(wines_data$Proline),
                                             mean(class_one$Proline),
                                             mean(class_two$Proline),
                                             mean(class_three$Proline)))
average_table
```

# e

Since there are three different classes, one will need to do 3 comparisons, class 1 vs. class 2, class 1 vs class 3 and class 2 vs class 3. Firstly, extracting the data of level of phenols of each classes:

```{r}
class_one_phenols <- class_one['Total_phenols']
class_two_phenols <- class_two['Total_phenols']
class_three_phenols <- class_three['Total_phenols']
```

For existing R function.

```{r}
t_test_1_2 <- t.test(class_one_phenols, class_two_phenols)
t_test_1_2

t_test_1_3 <- t.test(class_one_phenols, class_three_phenols)
t_test_1_3

t_test_2_3 <- t.test(class_two_phenols, class_three_phenols)
t_test_2_3
```

For manually conducting the t-test.

Then, calculating the mean, variance for each groups:

```{r}
mean_one <- mean(class_one_phenols[,])
mean_one
variance_one <- var(class_one_phenols[,])
variance_one

mean_two <- mean(class_two_phenols[,])
mean_two
variance_two <- var(class_two_phenols[,])
variance_two

mean_three <- mean(class_three_phenols[,])
mean_three
variance_three <- var(class_three_phenols[,])
variance_three
```

For different comparisons, assuming that the variances are different, first compute the t-statistics with formula: $t=\frac{(\hat{X}_1-\hat{X}_2) - (\mu_1-\mu_2)}{\sqrt{ \frac{S_1^2}{n_1} + \frac{S_2^2}{n_2}}}$, where $\hat{X}_1$ and $\hat{X}_2$ are the sample means, $\mu_1$ and $\mu_2$ are the means, $S_1^2$ and $S_2^2$ are the sample variances, $n_1$ and $n_2$ are the sizes.

Since the null hypothesis is that there is no difference between each class, $\mu_1-\mu_2=0$, thus the t-statistics are:

```{r}
t_1_2 <- (mean_one - mean_two) / 
  (sqrt((variance_one / num_class_one) + variance_two / num_class_two))
t_1_2 

t_1_3 <- (mean_one - mean_three) / 
  (sqrt((variance_one / num_class_one) + variance_three / num_class_three))
t_1_3

t_2_3 <- (mean_two - mean_three) / 
  (sqrt((variance_two / num_class_two) + variance_three / num_class_three))
t_2_3
```

Next the degrees of freedom are defined as $\nu=\frac{\left(\frac{S_1^2}{n_1} + \frac{S_2^2}{n_2}\right)^2}{\frac{\left(\frac{S_1^2}{n_1}\right)}{n_1-1} + \frac{\left(\frac{S_2^2}{n_2}\right)}{n_2-1}}$, then rounding it down to find the degree of freedom. The results are:

```{r}
nu_1_2 <- floor(
  (variance_one / num_class_one + variance_two / num_class_two)^2 / 
    ((variance_one / num_class_one)^2 / (num_class_one - 1) + 
       (variance_two / num_class_two)^2 / (num_class_two - 1))
  )
nu_1_2

nu_1_3 <- floor(
  (variance_one / num_class_one + variance_three / num_class_three)^2 / 
    ((variance_one / num_class_one)^2 / (num_class_one - 1) + 
       (variance_three / num_class_three)^2 / (num_class_three - 1))
  )
nu_1_3

nu_2_3 <- floor(
  (variance_two / num_class_two + variance_three / num_class_three)^2 / 
    ((variance_two / num_class_two)^2 / (num_class_two - 1) + 
       (variance_three / num_class_three)^2 / (num_class_three - 1))
  )
nu_2_3
```

Define a function to manually compute the p-value of give t-statistics and degree of freedom:

```{r}
compute_two_tail_p_value <- function(t_statistics, df){
  #inputs: t_statistics : the t-statistics, df: the degree of freedom
  #outputs: the p-value
  t_pdf <- function(x, df){
    # inputs x: variable, df: the degree of freedom
    # output the value of the probability density function
    return(gamma((df+1)/2) / (sqrt(df*pi) * gamma(df/2)) * (1 + (x^2)/df)^(-(df+1)/2))
  }

  p_value_two_tail <- 2 * integrate(t_pdf, t_statistics, Inf, df = df)$value
  
  return(p_value_two_tail)
}
p_1_2 <- compute_two_tail_p_value(t_1_2, nu_1_2)
p_1_2
p_1_3 <- compute_two_tail_p_value(t_1_3, nu_1_3)
p_1_3
p_2_3 <- compute_two_tail_p_value(t_2_3, nu_2_3)
p_2_3
```
Through calculation, one can observe that the p-values of all three comparisons are extremely small. Thus one can argue that there is extremely strong evidence against the null hypothesis for each pairwise comparison. The differences in phenol levels between all the classes are statistically significant.